Pre---
title: "FUT2 - Co-occurence network analysis"
output: html_document
author:
  - Samit Kundu
  - Gonçalo Correia
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
# Package and data import
```{r}
library(ggplot2)
library(readxl)
library(readr)
library(dplyr)
library(patchwork)
library(banocc)
library(igraph)
library(qgraph)
library(ggtext)
library(ggraph)
library(tidygraph)

microbiome_Species <- readxl::read_excel('./Data//FUT2-Dataset.xlsx', sheet='Species Level Dataset')
microbiome_ClinicalData <- readxl::read_excel('./Data//FUT2-Dataset.xlsx', sheet='Clinical Data')

FUT2Dataset <- merge(microbiome_ClinicalData, microbiome_Species, by.x='SampleID', by.y="SampleID")

colnames(FUT2Dataset)[colnames(FUT2Dataset) == 'Secretor Status'] <- "SecretorStatus"
FUT2Dataset$bmi <- as.numeric(FUT2Dataset$bmi)
FUT2Dataset$maternal.age <- as.numeric(FUT2Dataset$maternal.age)
FUT2Dataset$ABO <- factor(FUT2Dataset$ABO, levels=c('A', 'AB', 'B', 'O'))
FUT2Dataset$CST[grepl('IV', FUT2Dataset$CST)] <- 'IV'
FUT2Dataset$CST <- factor(FUT2Dataset$CST, levels=c('I', 'II', 'III', 'IV', 'V'))
FUT2Dataset$LactobacillusDominant <- factor(FUT2Dataset$LactobacillusDominant, levels=c('L. Dominant', 'L. Depleted'))
FUT2Dataset$LactobacillusDominant_CST <- factor(FUT2Dataset$LactobacillusDominant_CST, levels=c('L. Dominant', 'L. Depleted'))
```

Create network for the FUT2 microbiome data using the Banocc package
```{r}
# Select Timepoint 1 samples only
T1Samples <- subset(FUT2Dataset, Timepoint == 1)
T1Samples <- subset(T1Samples, !SampleID %in% c('T1A22', 'T1PG2'))

T1Samples_NSecretor <- subset(T1Samples, `SecretorStatus` == 'Non-secretor')
T1Samples_Secretor <- subset(T1Samples, `SecretorStatus` == 'Secretor')

# 37 is the first column with microbe counts
T1_SecretorCounts <- T1Samples_Secretor[, 36:ncol(T1Samples_Secretor)]
T1_NSecretorCounts <- T1Samples_NSecretor[, 36:ncol(T1Samples_NSecretor)]

# Subset only bacteria which are present in both secretors and non-secretors
presentBacteria <- names(which((colSums(T1_SecretorCounts > 0) > 2) & (colSums(T1_NSecretorCounts > 0) > 2)))
T1_SecretorCounts <- T1_SecretorCounts[, presentBacteria]
T1_NSecretorCounts <- T1_NSecretorCounts[, presentBacteria]

T1_Secretor_RelMatrix <- T1_SecretorCounts / rowSums(T1_SecretorCounts)
T1_Secretor_RelMatrix <- as.matrix(T1_Secretor_RelMatrix)
T1_NSecretor_RelMatrix <- T1_NSecretorCounts / rowSums(T1_NSecretorCounts)
T1_NSecretor_RelMatrix <- as.matrix(T1_NSecretor_RelMatrix)


compiled_banocc_model <- rstan::stan_model(model_code = banocc::banocc_model)

# Original iter = 10000
secretor_fit <- banocc::run_banocc(C = T1_Secretor_RelMatrix, compiled_banocc_model=compiled_banocc_model, 
                                   chains = 4, cores=4, verbose = T, iter = 1000, control = list(adapt_delta = 0.9))

# rstan::traceplot(secretor_fit$Fit, pars=paste0("O[1,", 2:9, "]"), inc_warmup=T)
secretor_out <- get_banocc_output(banoccfit = secretor_fit, conf_alpha = 0.05)
secretor_res <- secretor_out$Estimates.median

# Original iter = 10000
nsecretor_fit <- banocc::run_banocc(C = T1_NSecretor_RelMatrix, compiled_banocc_model=compiled_banocc_model, chains = 4, cores=4, verbose = T, iter = 1000, control = list(adapt_delta = 0.9))
# rstan::traceplot(nsecretor_fit$Fit, pars=paste0("O[1,", 2:9, "]"), inc_warmup=T)
nsecretor_out <- get_banocc_output(banoccfit = nsecretor_fit, conf_alpha = 0.05)
nsecretor_res <- nsecretor_out$Estimates.median
```
# Save/Load pre-calculated results
```{r}
#write.csv(nsecretor_res, file = "10K_NSecretor.csv")
#write.csv(secretor_res, file = "10K_Secretor.csv")

#saveRDS(secretor_fit, 'secretor_10K.R')
#saveRDS(nsecretor_fit, 'nsecretor_10K.R')

# secretor_res <- read_csv('./200K_Secretor.csv')
# nsecretor_res <- read_csv('./200K_NSecretor.csv')

secretor_res <- read.csv('./200K_Secretor.csv', header=T, row.names = 1)
secretor_res <- as.matrix(secretor_res)
colnames(secretor_res) <- rownames(secretor_res)

nsecretor_res <- read.csv('./200K_NSecretor.csv', header=T, row.names = 1)
nsecretor_res <- as.matrix(nsecretor_res)
colnames(nsecretor_res) <- rownames(nsecretor_res)

# secretor_fit <- readRDS('./secretor_200K.R')
# nsecretor_fit <- readRDS('./nsecretor_200K.R')

# nsecretor_out <- get_banocc_output(banoccfit = nsecretor_fit, conf_alpha = 0.05)
# nsecretor_res <- nsecretor_out$Estimates.median

# secretor_out <- get_banocc_output(banoccfit = secretor_fit, conf_alpha = 0.05)
# secretor_res <- secretor_out$Estimates.median
```

# Setup color labels
```{r}
create_label_list <- function(data) {
  list(
    `L. crispatus` = which(colnames(data) == 'Lactobacillus crispatus'),
    `L. iners` = which(colnames(data) == 'Lactobacillus iners'),
    `L. gasseri` = which(colnames(data) == 'Lactobacillus gasseri'),
    `L. jensenii` = which(colnames(data) == 'Lactobacillus jensenii'),
    `G. vaginalis` = which(colnames(data) %in% c('Gardnerella vaginalis', 'Gardnerella uncultured bacterium')),
    Lactobacillus = which(colnames(data) %in% c('Lactobacillus', 'Limosilactobacillus')),
    BV = which(colnames(data) %in% c('Atopobium vaginae', 'Prevotella timonensis', 'Aerococcus christensenii', 
                                     'Megasphaera unidentified', 'Prevotella bivia', 'Finegoldia magna', 
                                     'Prevotella colorans', 'Peptoniphilus uncultured bacterium', 'Prevotella buccalis', 
                                     'Anaerococcus sp.')),
    Pathobiont = which(colnames(data) %in% c('Streptococcus agalactiae', 'Streptococcus anginosus')),
    `B. breve` = which(colnames(data) == 'Bifidobacterium breve'),
    `Other/Unassigned` = which(colnames(data) == 'Domain Bacteria')
  )
}


create_node_colors <- function(label_list, node_names) {
  node_colors <- data.frame()
  for (x in names(label_list)) {
    for (y in label_list[[x]]) {
      node_colors <- rbind(node_colors, data.frame(label = y, Type = x))
    }
  }
  node_colors$label <- as.numeric(node_colors$label)
  node_colors$name <- node_names[node_colors$label]
  node_colors$Type <- factor(node_colors$Type, levels = c('L. crispatus', 'L. gasseri', 'L. iners', 'L. jensenii', 'G. vaginalis', 'Lactobacillus', 'BV', 'Pathobiont', 'B. breve', 'Other/Unassigned'))
  return(node_colors)
}

Label_List_Secretor <- create_label_list(secretor_res)
labelList_Secretor <- create_node_colors(Label_List_Secretor, rownames(secretor_res))

Label_List_NSecretor <- create_label_list(nsecretor_res)
labelList_NSecretor <- create_node_colors(Label_List_NSecretor,  rownames(nsecretor_res))
```

## Use qgraph and igraph for plotting and network statistics
```{r}
# qgraph::qgraph(as.matrix(secretor_res), layout="spring", vsize=2, border.width=1, groups=labelList_Secretor, color=c("#4daf4a","#ff7f00", "#377eb8","#ffff33","#a65628", "#984ea3", "#e41a1c", "#666666"),labels=F)

#qgraph::qgraph(nsecretor_res, layout="spring", vsize=2, border.width=1, groups=labelList_NSecretor, color=c("#4daf4a","#ff7f00", "#377eb8","#ffff33","#a65628", "#984ea3", "#e41a1c", "#666666"),labels=F)

# centralityPlot(secretor_res, include = c("Strength", "ExpectedInfluence","Closeness"), orderBy = "ExpectedInfluence")
# centralityPlot(nsecretor_res, include = c("Strength", "ExpectedInfluence","Closeness"), orderBy = "ExpectedInfluence")

nsec_cent <- centrality_auto(nsecretor_res)
nsec_cent_df <- as.data.frame(nsec_cent$node.centrality)
sec_cent <- centrality_auto(secretor_res)
sec_cent_df <- as.data.frame(sec_cent$node.centrality)
sec_cent_df$Taxon <- row.names(sec_cent_df)
nsec_cent_df$Taxon <- row.names(nsec_cent_df)

# ggplot() + geom_histogram(data=sec_cent_df, aes(x=ExpectedInfluence), fill="cyan", alpha=0.3) + geom_histogram(data=nsec_cent_df, aes(x=ExpectedInfluence), fill="red", alpha=0.3) + scale_fill_discrete() + theme_light() + scale_x_continuous(name="Expected Influence")

secretor_EIPlot <- ggplot(data=sec_cent_df, aes(x=Taxon, y=ExpectedInfluence)) + geom_bar(stat='identity', fill="#00BFC4", alpha=1)  + theme_light() + scale_y_continuous(name="Expected Influence") + coord_flip() + xlab(NULL) + theme(axis.text.y = element_blank())

nonsecretor_EIPlot <- ggplot(data=nsec_cent_df, aes(x=Taxon, y=ExpectedInfluence)) + geom_bar(stat='identity', fill="#F8766D", alpha=1)  + theme_light() + scale_y_continuous(name="Expected Influence") + coord_flip() + theme(axis.text.y=element_text(face=c("italic")))

EIPanel <- (nonsecretor_EIPlot | secretor_EIPlot)

ggsave(here::here('Figures', 'FigureS2.png'), EIPanel, dpi=300, width=10, height=8)
ggsave(here::here('Figures', 'FigureS2.svg'), EIPanel, dpi=300, width=10, height=8)


sec_qgrp <- qgraph::qgraph(secretor_res, layout="spring", vsize=4, border.width=1, groups=labelList_Secretor$Type, color=c("#4daf4a","#ff7f00", "#377eb8","#ffff33","#2134ca", "#a65628", "#984ea3", "#e41a1c", "plum2", "grey85"),labels=T)
nsec_qgrp <- qgraph::qgraph(nsecretor_res, layout="spring", vsize=4, border.width=1, groups=labelList_NSecretor$Type, color=c("#4daf4a","#ff7f00", "#377eb8","#ffff33","#2134ca", "#a65628", "#984ea3", "#e41a1c", "plum2",  "grey85"), labels=T)
```

# Non-secretor network plot
```{r}
CustomColorsValues=c("L. crispatus"="#4daf4a", "L. iners"="#ff7f00", "L. gasseri"="#377eb8", "G. vaginalis"="deeppink", 
                     "L. jensenii"="#ffff33", "Lactobacillus"="#a65628", "BV"="#984ea3", "Pathobiont"="#e41a1c", "B. breve"="plum2", "Other/Unassigned" = "grey75")

# plot using ggraph
graph_tbl_nsec <- nsec_qgrp %>% 
  as_tbl_graph() %>% 
  activate(nodes) %>% left_join(labelList_NSecretor, by = join_by(label == label))

# use the default laytouts from igraph
layout_nsec <- create_layout(graph_tbl_nsec, layout = 'manual', x=nsec_qgrp$layout[, 1], y=nsec_qgrp$layout[,2 ])

# layout <- create_layout(graph_tbl, layout = 'igraph', algorithm='star', center=9)
networkNonSecretor <- ggraph(layout_nsec) +
  geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F) + 
  geom_node_circle(aes(fill = Type, r=0.075, color=Type),  show.legend = T) + 
  geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=4) + 
  scale_fill_manual(values=CustomColorsValues) + scale_colour_manual(values=CustomColorsValues) + 
  scale_edge_colour_gradient2(low = "firebrick3",mid = "white", high = "darkgreen", midpoint = 0, 
  limits = c(-0.45,0.8), space = "Lab", na.value = "grey50",guide = "edge_colourbar", aesthetics = "edge_colour") + 
  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + 
  theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif"))

```

# Secretor network plot
```{r}
CustomColorsValues=c("L. crispatus"="#4daf4a", "L. iners"="#ff7f00", "L. gasseri"="#377eb8", "G. vaginalis"="deeppink", 
                     "L. jensenii"="#ffff33", "Lactobacillus"="#a65628", "BV"="#984ea3", "Pathobiont"="#e41a1c", "B. breve"="plum2", "Other/Unassigned" = "grey75")

# plot using ggraph
graph_tbl_sec <- sec_qgrp %>% 
  as_tbl_graph() %>% 
  activate(nodes) %>% left_join(labelList_Secretor, by = join_by(label == label))

# use the default laytouts from igraph
layout_sec <- create_layout(graph_tbl_sec, layout = 'manual', x=sec_qgrp$layout[, 1], y=sec_qgrp$layout[,2 ])

# layout <- create_layout(graph_tbl, layout = 'igraph', algorithm='star', center=9)
networkSecretor <- ggraph(layout_sec) +
   geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +  geom_node_circle(aes(fill = Type, r=0.075, color=Type),  show.legend = T)  + 
  geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=4) +
 scale_fill_manual(values=CustomColorsValues)  +  scale_colour_manual(values=CustomColorsValues)  +
  scale_edge_colour_gradient2(low = "firebrick3", mid = "white", high = "darkgreen", midpoint = 0, limits = c(-0.45,0.8), 
  space = "Lab", na.value = "grey50",guide = "edge_colourbar", aesthetics = "edge_colour") + 
  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + 
  theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif"))

networkSecretor <- networkSecretor + ggtitle('Secretor') +
    theme(plot.title = element_text(hjust = 0.5, size = 30, face = "plain"))

networkNonSecretor <- networkNonSecretor + ggtitle('Non-secretor') +
    theme(plot.title = element_text(hjust = 0.5, size = 30, face = "plain"))
networkPanel <- (networkSecretor | networkNonSecretor )  & theme(plot.margin =  unit(c(0,0,0,0), 'cm'))
networkPanel <- networkPanel & theme(plot.tag = element_text(size=35))

ggsave(here::here('Figures', 'Figure2_TopNetworks.png'), dpi=300, width=12, height=6)
```

##Flow diagrams
These re-center the network on specific nodes and are useful for highlighting the distribution of edges.
```{r}
par(mfrow=c(2,1))
#flow diagrams for L.crispatus
bugIdx <- which(colnames(secretor_res) == 'Lactobacillus crispatus')
flow(sec_qgrp, bugIdx, vsize=3)
flow(nsec_qgrp, bugIdx, vsize=3)
#flow diagram for L.gasseri
bugIdx <- which(colnames(secretor_res) == 'Lactobacillus gasseri')
flow(sec_qgrp, bugIdx, vsize=3)
flow(nsec_qgrp, bugIdx, vsize=3)
#flow diagram for L.iners
bugIdx <- which(colnames(secretor_res) == 'Lactobacillus iners')
flow(sec_qgrp, bugIdx, vsize=3)
flow(nsec_qgrp, bugIdx, vsize=3)
#flow diagram for L.jensenii
bugIdx <- which(colnames(secretor_res) == 'Lactobacillus jensenii')
flow(sec_qgrp, bugIdx, vsize=3)
flow(nsec_qgrp, bugIdx, vsize=3)
#flow diagram for G. vaginalis
bugIdx <- which(colnames(secretor_res) == 'Gardnerella vaginalis')
flow(sec_qgrp, bugIdx, vsize=3)
flow(nsec_qgrp, bugIdx, vsize=3)
```


# Pairwise star plots
# L. crispatus
```{r}
bugIdx <- which(colnames(secretor_res) == 'Lactobacillus crispatus')

# use the default laytouts from igraph
secSubGraph_LCrisp <- graph_tbl_sec %>% activate(edges) %>% filter(from== bugIdx | to == bugIdx)

layout_sec_crisp <- create_layout(secSubGraph_LCrisp, layout = 'igraph', algorithm='star', center=bugIdx)

starSec_LCrisp <- ggraph(layout_sec_crisp) +
  geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +
  geom_node_point(
    aes(color = Type), size = 10,
    show.legend = T
  ) + scale_color_manual(values=CustomColorsValues)  + 
#  geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=5)  + 
  scale_edge_colour_gradient2(low = "firebrick3",
  mid = "white",high = "darkgreen", midpoint = 0,  limits = c(-0.45,0.8), space = "Lab", na.value = "grey50",guide = "none",
  aesthetics = "edge_colour") +  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif"))  +  theme(plot.margin = margin(0, 0, 0, 0, "pt")) + guides(colour="none")

nsecSubGraph_LCrisp <- graph_tbl_nsec %>% activate(edges) %>% filter(from==bugIdx | to == bugIdx)
layout_Nsec_crisp <- create_layout(nsecSubGraph_LCrisp, layout = 'igraph', algorithm='star', center=bugIdx)

starNSec_LCrisp <- ggraph(layout_Nsec_crisp) +
  geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +
  geom_node_point(
    aes(color = Type), size = 10,
    show.legend = T
  ) + scale_color_manual(values=CustomColorsValues)  + 
#  geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=5)  + 
  scale_edge_colour_gradient2(low = "firebrick3",
  mid = "white",high = "darkgreen", midpoint = 0,  limits = c(-0.45,0.8), space = "Lab", na.value = "grey50",guide = "none",
  aesthetics = "edge_colour") +  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif")) +  theme(plot.margin = margin(0, 0, 0, 0, "pt")) + guides(colour="none")

starDiagram_Lcrisp <- (starSec_LCrisp | starNSec_LCrisp) # +  plot_layout(guides = 'collect')
```

# L. iners
```{r}
bugIdx <- which(colnames(secretor_res) == 'Lactobacillus iners')

# use the default laytouts from igraph
secSubGraph_Liners <- graph_tbl_sec %>% activate(edges) %>% filter(from == bugIdx | to == bugIdx)

layout_sec_liners <- create_layout(secSubGraph_Liners, layout = 'igraph', algorithm='star', center=bugIdx)

starSec_Liners <- ggraph(layout_sec_liners) +
 geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +
  geom_node_point(
    aes(color = Type), size = 10,
    show.legend = T
  ) + scale_color_manual(values=CustomColorsValues) + 
#  geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=5)  + 
  scale_edge_colour_gradient2(low = "firebrick3",
  mid = "white",high = "darkgreen", midpoint = 0, limits = c(-0.45,0.8),  space = "Lab", na.value = "grey50",guide = "none",
  aesthetics = "edge_colour") +  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif"))  +  theme(plot.margin = margin(0, 0, 0, 0, "pt")) + guides(colour="none")

nsecSubGraph_Liners <- graph_tbl_nsec %>% activate(edges) %>% filter(from==bugIdx | to == bugIdx)
layout_Nsec_liners <- create_layout(nsecSubGraph_Liners, layout = 'igraph', algorithm='star', center=bugIdx)

starNSec_Liners <- ggraph(layout_Nsec_liners) +
  geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +
  geom_node_point(
    aes(color = Type), size = 10,
    show.legend = T
  ) + scale_color_manual(values=CustomColorsValues) + 
#  geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=5)  + 
  scale_edge_colour_gradient2(low = "firebrick3",
  mid = "white",high = "darkgreen", midpoint = 0, limits = c(-0.45,0.8),  space = "Lab", na.value = "grey50",guide = "none",
  aesthetics = "edge_colour") +  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif")) +  theme(plot.margin = margin(0, 0, 0, 0, "pt")) + guides(colour="none")

starDiagram_Liners <- (starSec_Liners | starNSec_Liners) # +  plot_layout(guides = 'collect')
```

# L. gasseri
```{r}
bugIdx <- which(colnames(secretor_res) == 'Lactobacillus gasseri')

# use the default laytouts from igraph
secSubGraph_Lgasseri <- graph_tbl_sec %>% activate(edges) %>% filter(from == bugIdx | to == bugIdx)

layout_sec_lgasseri <- create_layout(secSubGraph_Lgasseri, layout = 'igraph', algorithm='star', center=bugIdx)

starSec_Lgasseri <- ggraph(layout_sec_lgasseri) +
  geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +
  geom_node_point(
    aes(color = Type), size = 10,
    show.legend = T
  ) + scale_color_manual(values=CustomColorsValues)  + 
 # geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=5)  +
  scale_edge_colour_gradient2(low = "firebrick3",
  mid = "white",high = "darkgreen", midpoint = 0, limits = c(-0.45,0.8),  space = "Lab", na.value = "grey50",guide = "none",
  aesthetics = "edge_colour") +  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif"))  +  theme(plot.margin = margin(0, 0, 0, 0, "pt")) + guides(colour="none")

nsecSubGraph_Lgasseri <- graph_tbl_nsec %>% activate(edges) %>% filter(from == bugIdx | to == bugIdx)
layout_Nsec_lgasseri <- create_layout(nsecSubGraph_Lgasseri, layout = 'igraph', algorithm='star', center=bugIdx)

starNSec_Lgasseri <- ggraph(layout_Nsec_lgasseri) +
  geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +
  geom_node_point(
    aes(color = Type), size = 10,
    show.legend = T
  ) + scale_color_manual(values=CustomColorsValues)  + 
  # geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=5) + 
  scale_edge_colour_gradient2(low = "firebrick3", mid = "white",high = "darkgreen", midpoint = 0, limits = c(-0.45,0.8),  
  space = "Lab", na.value = "grey50",guide = "none", aesthetics = "edge_colour") +  
  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif")) +  theme(plot.margin = margin(0, 0, 0, 0, "pt")) + guides(colour="none")

starDiagram_Lgasseri <- (starSec_Lgasseri | starNSec_Lgasseri)#  +  plot_layout(guides = 'collect')
```

# L. jensenii
```{r}
bugIdx <- which(colnames(secretor_res) == 'Lactobacillus jensenii')

# use the default laytouts from igraph
secSubGraph_Ljenseni <- graph_tbl_sec %>% activate(edges) %>% filter(from == bugIdx | to == bugIdx)

layout_sec_ljenseni <- create_layout(secSubGraph_Ljenseni, layout = 'igraph', algorithm='star', center=bugIdx)

starSec_Ljenseni <- ggraph(layout_sec_ljenseni) +
 geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +
  geom_node_point(
    aes(color = Type), size = 10,
    show.legend = T
  ) + scale_color_manual(values=CustomColorsValues)  + 
  # geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=5)  + 
  scale_edge_colour_gradient2(low = "firebrick3",
  mid = "white",high = "darkgreen", midpoint = 0, limits = c(-0.45,0.8),  space = "Lab", na.value = "grey50",guide = "none",
  aesthetics = "edge_colour") +  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif"))  +  theme(plot.margin = margin(0, 0, 0, 0, "pt")) + guides(colour="none")

nsecSubGraph_Ljenseni <- graph_tbl_nsec %>% activate(edges) %>% filter(from==bugIdx | to == bugIdx)
layout_Nsec_ljenseni <- create_layout(nsecSubGraph_Ljenseni, layout = 'igraph', algorithm='star', center=bugIdx)

starNSec_Ljenseni <- ggraph(layout_Nsec_ljenseni) +
 geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +
  geom_node_point(
    aes(color = Type), size = 10,
    show.legend = T
  ) + scale_color_manual(values=CustomColorsValues)  + 
  # geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=5) + 
  scale_edge_colour_gradient2(low = "firebrick3",
  mid = "white",high = "darkgreen", midpoint = 0, limits = c(-0.45,0.8),  space = "Lab", na.value = "grey50",guide = "none",
  aesthetics = "edge_colour") +  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif")) +  theme(plot.margin = margin(0, 0, 0, 0, "pt")) + guides(colour="none")

starDiagram_Ljenseni <- (starSec_Ljenseni | starNSec_Ljenseni) # +  plot_layout(guides = 'collect')
```

# G. vaginalis
```{r}
# use the default laytouts from igraph
bugIdx <- which(colnames(secretor_res) == 'Gardnerella vaginalis')

secSubGraph_Gvaginalis <- graph_tbl_sec %>% activate(edges) %>% filter(from == bugIdx | to == bugIdx)

layout_sec_gvaginalis <- create_layout(secSubGraph_Gvaginalis, layout = 'igraph', algorithm='star', center= bugIdx)

starSec_Gvaginalis <- ggraph(layout_sec_gvaginalis) +
  geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +
  geom_node_point(
    aes(color = Type), size = 10,
    show.legend = T
  ) + scale_color_manual(values=CustomColorsValues)  + 
  # geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=5) + 
  scale_edge_colour_gradient2(low = "firebrick3",
  mid = "white",high = "darkgreen", midpoint = 0, limits = c(-0.45,0.8),  space = "Lab", na.value = "grey50", guide = "none",
  aesthetics = "edge_colour") +  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif"))  +  theme(plot.margin = margin(0, 0, 0, 0, "pt")) + guides(colour="none")

nsecSubGraph_Gvaginalis <- graph_tbl_nsec %>% activate(edges) %>% filter(from == bugIdx | to == bugIdx)
layout_Nsec_gvaginalis <- create_layout(nsecSubGraph_Gvaginalis, layout = 'igraph', algorithm='star', center = bugIdx)

starNSec_Gvaginalis <- ggraph(layout_Nsec_gvaginalis) +
  geom_edge_fan(
    aes(color = weight, edge_linewidth=abs(weight)),
    show.legend = F
  ) +
  geom_node_point(
    aes(color = Type), size = 10,
    show.legend = T
  ) + scale_color_manual(values=CustomColorsValues)  + 
  # geom_node_text(aes(label = name), col='gray0', repel = T, fontface=3, size=5) + 
  scale_edge_colour_gradient2(low = "firebrick3",
  mid = "white",high = "darkgreen", midpoint = 0, limits = c(-0.45,0.8),  space = "Lab", na.value = "grey50", guide = "none",
  aesthetics = "edge_colour") +  theme_graph() + scale_edge_width(range = c(0, 2), guide = "none") + theme(legend.text = element_text(size = 10,  family="sans-serif"), legend.title=element_text(size=12,  family="sans-serif")) +  theme(plot.margin = margin(0, 0, 0, 0, "pt")) + guides(colour="none")

starDiagram_Gvaginalis <- (starSec_Gvaginalis | starNSec_Gvaginalis) #+  plot_layout(guides = 'collect')
```

```{r}
starDiagramFigures <- (starDiagram_Lcrisp / starDiagram_Liners /   starDiagram_Gvaginalis) 

# starDiagramFigures <- (starDiagram_Lcrisp / starDiagram_Liners / starDiagram_Lgasseri / starDiagram_Ljenseni / starDiagram_Gvaginalis) 

starDiagramFigures <- starDiagramFigures + plot_layout(guides='collect')

# starDiagramFigures <- starDiagramFigures + plot_annotation(tag_levels = list(c('A', '', 'B', '', 'C', '', 'D', '', 'E')))

starDiagramFigures <- starDiagramFigures
networkFigure <- (networkPanel / starDiagramFigures) 
networkFigure <- networkFigure + plot_layout(heights = c(1, 2), guides='collect')

networkFigure <- networkFigure + plot_annotation(tag_levels = list(c('A', '', 'B', '', 'C', '', 'D', '')))
networkFigure <- networkFigure & theme(plot.tag = element_text(size=35))

ggsave(here::here('Figures', 'Fig2_NetworkPanel.png'), networkFigure, dpi=300, height=20, width=14)
ggsave(here::here('Figures', 'Fig2_NetworkPanel.svg'), networkFigure, dpi=300, height=20, width=14)
```
