---
title: "FUT2 Project - Microbiome analyses"
output: html_document
author:
  - Samit Kundu
  - Gonçalo Correia
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Analyses to perform:
### Overview of differences between secretors and non-secretors with diversity measures and Principal component analysis
Comparison of observed species, Shannon alpha-diversity, and dispersion of principal components 
### Differential abundance analysis
Perform differential abundance analysis on the FUT2 microbiome species-level data
The analysis will be done on the first two timepoints only (Early and Mid) as by the third timepoint some patients might have delivered
1. Comparison of the microbial composition between secretors and non-secretors at timepoint 1 (Early).
2. Comparison of the microbial composition between secretors and non-secretors at timepoint 2 (Mid).

# Package and data import
```{r}
library(ggplot2)
library(ALDEx2)
library(readxl)
library(readr)
library(dplyr)
library(zCompositions)
library(ggstatsplot)
library(gginnards)
library(patchwork)
library(reshape2)
library(ggtext)
library(vegan)
library(gt)
library(gtsummary)

microbiome_Species <- readxl::read_excel('./Data//FUT2-Dataset.xlsx', sheet='Species Level Dataset')
microbiome_Genera <- readxl::read_excel('./Data//FUT2-Dataset.xlsx', sheet='Genera Level Dataset')
microbiome_ClinicalData <- readxl::read_excel('./Data//FUT2-Dataset.xlsx', sheet='Clinical Data')
```

```{r}
FUT2Dataset <- merge(microbiome_ClinicalData, microbiome_Species, by.x='SampleID', by.y="SampleID")

FUT2Dataset_Genera_RelativeAbundances <-  merge(microbiome_ClinicalData, microbiome_Genera, by.x='SampleID', by.y="SampleID")

colnames(FUT2Dataset)[colnames(FUT2Dataset) == 'Secretor Status'] <- "SecretorStatus"
FUT2Dataset$bmi <- as.numeric(FUT2Dataset$bmi)
FUT2Dataset$maternal.age <- as.numeric(FUT2Dataset$maternal.age)
FUT2Dataset$ABO <- factor(FUT2Dataset$ABO, levels=c('A', 'AB', 'B', 'O'))
FUT2Dataset$CST[grepl('IV', FUT2Dataset$CST)] <- 'IV'
FUT2Dataset$CST <- factor(FUT2Dataset$CST, levels=c('I', 'II', 'III', 'IV', 'V'))
FUT2Dataset$LactobacillusDominant <- factor(FUT2Dataset$LactobacillusDominant, levels=c('L. Dominant', 'L. Depleted'))
FUT2Dataset$LactobacillusDominant_CST <- factor(FUT2Dataset$LactobacillusDominant_CST, levels=c('L. Dominant', 'L. Depleted'))
```
# Figure 1 (A)
```{r}
taxaIdx <- 36:ncol(FUT2Dataset_Genera_RelativeAbundances)
# Add 1 extra to account for "Kingdom.Bacteria" Taxa
top10Taxa <- colnames(FUT2Dataset_Genera_RelativeAbundances[, taxaIdx][order(colSums(FUT2Dataset_Genera_RelativeAbundances[, taxaIdx]), decreasing=T)])[1:11]
# Remove unassigned reads
top10Taxa <- top10Taxa[c(-7)]

top10Taxa <- sort(as.character(top10Taxa))
plotDataset <- FUT2Dataset_Genera_RelativeAbundances[, c('Timepoint', 'Secretor Status', top10Taxa)]

plotDataset <- melt(plotDataset, id.vars=c('Timepoint', 'Secretor Status'))

relabundanceGenera_TP1 <- ggplot(subset(plotDataset, Timepoint==1), aes(x=value, y=variable, color=`Secretor Status`)) + geom_jitter(position=position_jitterdodge()) + theme_minimal() + xlab('Relative abundance') + ylab(NULL) + theme(axis.text.y = element_text(size=12, face = "italic"), legend.position = 'none', axis.text.x = element_text(size=12), axis.title.x = element_text(size=14, face="bold"), axis.title.y = element_text(size=14,  face="bold"))

relabundanceGenera_TP2 <- ggplot(subset(plotDataset, Timepoint==2), aes(x=value, y=variable, color=`Secretor Status`)) + geom_jitter(position=position_jitterdodge()) + theme_minimal() + ylab(NULL) + xlab('Relative abundance')  + theme(axis.text.y = element_text(size=12, face = "italic"), legend.position = 'none', axis.text.x = element_text(size=12), axis.title.x = element_text(size=14, face="bold"), axis.title.y = element_text(size=14,  face="bold"))

relabundanceGenera_TP3 <- ggplot(subset(plotDataset, Timepoint==3), aes(x=value, y=variable, color=`Secretor Status`)) + geom_jitter(position=position_jitterdodge()) + theme_minimal() + ylab(NULL) + xlab('Relative abundance') + theme(axis.text.y = element_text(size=12, face = "italic"), legend.position = 'none', axis.text.x = element_text(size=12), axis.title.x = element_text(size=14, face="bold"), axis.title.y = element_text(size=14,  face="bold"))

relativeAbundanceGenera_Panel <- (relabundanceGenera_TP1 | relabundanceGenera_TP2 | relabundanceGenera_TP3) + plot_annotation(tag_levels = list(c('A', '', '')))
relativeAbundanceGenera_Panel <- relativeAbundanceGenera_Panel + plot_layout(guides='collect') & xlab(NULL)  #+ plot_annotation(tag_levels = list(c('Early', 'Mid', 'Late'))) &  theme(plot.tag.position = 'top') & xlab(NULL)  

```

# Figure 1 (B)
```{r}
T1_Counts <- table(subset(FUT2Dataset_Genera_RelativeAbundances, Timepoint==1)$`Secretor Status`)
LactoRelAbundance_TP1 <- ggplot(subset(FUT2Dataset_Genera_RelativeAbundances, Timepoint==1), aes(x=`Secretor Status`, y=Lactobacillus, fill=`Secretor Status`, color=`Secretor Status`)) + geom_jitter(position=position_jitterdodge()) + theme_minimal() +  ylab('*Lactobacillus* relative abundance') + xlab(NULL) +  scale_x_discrete(labels=apply(T1_Counts, 1, function(x){paste0('n=', x)})) + theme(axis.text.y = element_text(size=12, face = "italic"), axis.text.x = element_text(size=12), axis.title.x = element_text(size=14, face="bold"), axis.title.y = element_markdown(size=14,  face="bold"))

T2_Counts <- table(subset(FUT2Dataset_Genera_RelativeAbundances, Timepoint==2)$`Secretor Status`)
LactoRelAbundance_TP2 <- ggplot(subset(FUT2Dataset_Genera_RelativeAbundances, Timepoint==2), aes(x=`Secretor Status`, y=Lactobacillus, fill=`Secretor Status`, color=`Secretor Status`)) + geom_jitter(position=position_jitterdodge()) + theme_minimal()  + xlab(NULL) + ylab(NULL) + scale_x_discrete(labels=apply(T2_Counts, 1, function(x){paste0('n=', x)})) + theme(axis.text.y = element_text(size=12, face = "italic"), axis.text.x = element_text(size=12), axis.title.x = element_text(size=14, face="bold"), axis.title.y = element_markdown(size=14,  face="bold"))

T3_Counts <- table(subset(FUT2Dataset_Genera_RelativeAbundances, Timepoint==3)$`Secretor Status`)
LactoRelAbundance_TP3 <- ggplot(subset(FUT2Dataset_Genera_RelativeAbundances, Timepoint==3), aes(x=`Secretor Status`, y=Lactobacillus, fill=`Secretor Status`, color=`Secretor Status`)) + geom_jitter(position=position_jitterdodge()) + theme_minimal() + xlab(NULL) +  ylab(NULL) + scale_x_discrete(labels=apply(T3_Counts, 1, function(x){paste0('n=', x)}))  + theme(axis.text.y = element_text(size=12, face = "italic"), axis.text.x = element_text(size=12), axis.title.x = element_text(size=14, face="bold"), axis.title.y = element_markdown(size=14,  face="bold"))

relativeAbundanceLactobacillus_Panel <- (LactoRelAbundance_TP1 | LactoRelAbundance_TP2 | LactoRelAbundance_TP3) + plot_layout(guides='collect') + plot_annotation(tag_levels = list(c('B', '', ''))) &  theme(legend.position='bottom')

```


```{r}
CustomColours <- c('I'='forestgreen', 'II'='steelblue3', 'III'='darkorange3', 'IV'='firebrick3', 'V'='turquoise4')

SecretorCST_1 <- ggbarstats(subset(FUT2Dataset, Timepoint == 1), x=CST, y=SecretorStatus, results.subtitle=F) 
SecretorCST_1 <- delete_layers(SecretorCST_1, "GeomLabel")
SecretorCST_1 <- SecretorCST_1 + scale_fill_manual(values=CustomColours) + theme(legend.title = element_text(size=12), axis.text.x = element_text(size = 10), axis.title.x=element_text(size=12), axis.title.y=element_text(size=12))   + theme(axis.text.x = element_text(angle = 0,  size = 12)) + xlab(NULL)


SecretorCST_2 <- ggbarstats(subset(FUT2Dataset, Timepoint == 2), x=CST, y=SecretorStatus, results.subtitle=F) 
SecretorCST_2 <- delete_layers(SecretorCST_2, "GeomLabel")
SecretorCST_2 <- SecretorCST_2 + scale_fill_manual(values=CustomColours) + theme(legend.title = element_text(size=12), axis.text.x = element_text(size = 10), axis.title.x=element_text(size=12), axis.title.y=element_text(size=12))   + theme(axis.text.x = element_text(angle = 0,  size = 12)) + xlab(NULL)


SecretorCST_3 <- ggbarstats(subset(FUT2Dataset, Timepoint == 3), x=CST, y=SecretorStatus, results.subtitle=F) 
SecretorCST_3 <- delete_layers(SecretorCST_3, "GeomLabel")
SecretorCST_3 <- SecretorCST_3 + scale_fill_manual(values=CustomColours) + theme(legend.title = element_text(size=12), axis.text.x = element_text(size = 10), axis.title.x=element_text(size=12), axis.title.y=element_text(size=12)) + theme(axis.text.x = element_text(angle = 0, size = 12)) + xlab(NULL)

CST_SecretorStatusPanel <- (SecretorCST_1 | SecretorCST_2 | SecretorCST_3) #
CST_SecretorStatusPanel <- CST_SecretorStatusPanel + plot_layout(guides = "collect") + plot_annotation(tag_levels = list(c('C', '', '')))
```

# Assemble Figure 1
```{r}
Figure1Panel <- (relativeAbundanceGenera_Panel / relativeAbundanceLactobacillus_Panel / CST_SecretorStatusPanel) + 
  plot_annotation(tag_levels =list(c('A', '', '', 'B', '', '', 'C', '', '')))  # & theme(legend.justification = "left")

Figure1Panel <- Figure1Panel & theme(plot.tag = element_text(size=25))

ggsave(here::here('Figures', 'Figure1.png'), Figure1Panel, dpi=300, height=14, width=10)
ggsave(here::here('Figures', 'Figure1.svg'), Figure1Panel, dpi=300, height=14, width=10)

```

# CST distribution per timepoint summary
```{r}
CST_Distribution <- FUT2Dataset %>% select(Timepoint, CST) %>% group_by(Timepoint, CST) %>% summarise(n=n()) %>% mutate(freq = format((n / sum(n))*100, nsmall=2, digits=2)) %>% gt()
```

# Figure S1 - Diversity boxplots
```{r}
# Timepoint 1 "Early"
TP1Samples <- subset(FUT2Dataset, Timepoint == 1)
divPlotFrame <- data.frame(cbind(SecretorStatus=TP1Samples$SecretorStatus, 
                                 Nobs=as.numeric(rowSums(TP1Samples[, 36:ncol(TP1Samples)] > 0))), 
                           alpha_species=diversity(TP1Samples[, 36:ncol(TP1Samples)]))


nObsPlot <- ggplot(divPlotFrame, aes(x=SecretorStatus, y=as.numeric(Nobs))) + geom_boxplot(outlier.shape=NA) + geom_jitter(aes(col=SecretorStatus)) + theme_light() + ylab('Observed species') + guides(color=guide_legend(title="Secretor Status")) + xlab(NULL) + theme(axis.text.y = element_text(size=12), axis.text.x = element_text(size=12), axis.title.x = element_text(size=14), axis.title.y = element_markdown(size=14))


shannonPlot <- ggplot(divPlotFrame, aes(x=SecretorStatus, y=as.numeric(alpha_species))) + geom_boxplot(outlier.shape=NA) + geom_jitter(aes(col=SecretorStatus)) + theme_light() + ylab('Shannon') + guides(color=guide_legend(title="Secretor Status")) + xlab(NULL) + theme(axis.text.y = element_text(size=12), axis.text.x = element_text(size=12), axis.title.x = element_text(size=14), axis.title.y = element_markdown(size=14))

FigureS1Div <- (nObsPlot | shannonPlot) + plot_layout(guides='collect')

wilcox.test(as.numeric(Nobs) ~ SecretorStatus, data=divPlotFrame)

wilcox.test(as.numeric(alpha_species) ~ SecretorStatus, data=divPlotFrame)


```

# Figure S1 - PCA plots
## CLR transform counts before  PCA
```{r}
# Timepoint 1 "Early"
T1Samples <- subset(FUT2Dataset, Timepoint == 1)

T1Counts <- T1Samples[, 36:ncol(T1Samples)]
# Exclude taxa without counts in TP1
excludeTaxa <- which(colSums(T1Counts) == 0)

if (length(excludeTaxa) > 0) {
T1Counts <- subset(T1Counts, select=-c(as.numeric(excludeTaxa)))
}

T1Samples.czm <- cmultRepl(T1Counts, label=0, method = "CZM", output="p-counts", z.warning = 1)

T1Samples.clr <- apply(T1Samples.czm, 1, function(x){log(x) - mean(log(x))})
T1Samples.clr <- t(T1Samples.clr)
```

# PCA model
```{r}
pcaModel <- pcaMethods::pca(T1Samples.clr, nPcs = 6, scale = 'none')

PCAFrame <- data.frame(cbind(Secretor=T1Samples$SecretorStatus,  pcaModel@scores))
PCAFrame[, 2:ncol(PCAFrame)] <- apply(PCAFrame[, 2:ncol(PCAFrame)], 2, as.numeric)

PCAScores_12 <- ggplot(PCAFrame, aes(x=PC1, y=PC2)) + geom_point(aes(col=Secretor)) + theme_light() + stat_ellipse(aes(group=Secretor, col=Secretor)) + xlab(paste0('PC1: ', format(pcaModel@R2[1]*100, nsmall=2, digits=3), ' %')) +  ylab(paste0('PC2: ', format(pcaModel@R2[2]*100, nsmall=2, digits=3), ' %')) + theme(axis.text.y = element_text(size=12), axis.text.x = element_text(size=12), axis.title.x = element_text(size=14), axis.title.y = element_markdown(size=14))
PCAScores_34 <- ggplot(PCAFrame, aes(x=PC3, y=PC4)) + geom_point(aes(col=Secretor)) + theme_light() + stat_ellipse(aes(group=Secretor, col=Secretor)) + xlab(paste0('PC3: ', format(pcaModel@R2[3]*100, nsmall=2, digits=3), ' %')) +  ylab(paste0('PC4: ', format(pcaModel@R2[4]*100, nsmall=2, digits=3), ' %'))  + theme(axis.text.y = element_text(size=12), axis.text.x = element_text(size=12), axis.title.x = element_text(size=14), axis.title.y = element_markdown(size=14))


PCAScores <- (PCAScores_12 | PCAScores_34) + plot_layout(guides='collect')
```

```{r}
permanovaModel <- adonis2(PCAFrame[, 2:ncol(PCAFrame)] ~ Secretor, data = PCAFrame, method='euclidean', permutations=1000, parallel=6, by='margin')

results_table <-
  permanovaModel %>%
  gt(rownames_to_stub = TRUE) %>%
  fmt_number(
    columns = c(4,5,6),
    decimals = 3,
    suffixing = F
  ) %>%
  fmt_missing(columns = c(5,6), 
              missing_text = '')
gtsave(results_table, here::here('Figures', 'T1_PERMANOVA.html'))

```


# Analysis of early pregnancy sample
```{r}
T1Samples <- subset(FUT2Dataset, Timepoint == 1)

# 27 when everything is merged as is, double check
t1.aldex <- aldex.clr(t(T1Samples[, 36:dim(T1Samples)[2]]), T1Samples$`SecretorStatus`, mc.samples=1000, verbose=T)
t1.aldex.tt <- aldex.ttest(t1.aldex, paired.test=F, verbose=T)
t1.aldex.effect <- aldex.effect(t1.aldex, verbose=F)
t1.all <- data.frame(t1.aldex.tt, t1.aldex.effect, stringsAsFactors = F)
```
### Comparison of secretors and non-secretors at Early pregnancy
No individual organisms are significantly differentially abundant and the effect sizes are very small.

Plot as horizontal barplot in order of effect size
```{r}
par(mfrow=c(1,2))
aldex.plot(t1.all, type="MA", test="welch")

aldex.plot(t1.all, type="MW", test="welch")
t1.all$spp <- with(t1.all, reorder(rownames(t1.all), effect))
#t1.all.order <- t1.all[order(t1.all$effect),]
t1.all[["sign"]] = ifelse(t1.all[["effect"]] >= 0, "Secretor", "Non-secretor")

par(mfrow=c(1,1))
t1.all_topN <- t1.all[order(abs(t1.all$effect), decreasing=T), ][1:20, ]
t1.effect.plot <- ggplot(t1.all_topN, aes(x=spp, y=effect, fill=sign)) + geom_bar(stat="identity", width=0.5) + coord_flip() + theme_light() + scale_fill_discrete(name="Enriched in") + ylab('Effect') + xlab('Taxon') + theme(axis.text.y=element_markdown(size=12, face=c("italic")), axis.text.x = element_text(size=12), axis.title.x = element_text(size=14), axis.title.y = element_markdown(size=14))
t1.effect.plot
```

# Figure S1 assembled
```{r}
FigureS1 <- (FigureS1Div / PCAScores) # + plot_annotation(tag_levels=list(c('A', '', 'B', '')))
t1.effect.plot <- t1.effect.plot # +  plot_annotation(tag_levels=list(c('C')))

FigureS1 <- (FigureS1 / t1.effect.plot) +  plot_annotation(tag_levels=list(c('A', '', 'B', '', 'C'))) #  + plot_layout(heights = c(3, 1))
FigureS1[[1]][[1]] <- FigureS1[[1]][[1]] + theme(axis.title.y = element_markdown(vjust = -35))
FigureS1[[2]][[1]] <- FigureS1[[2]][[1]] + theme(axis.title.y = element_markdown(vjust = -35))
# FigureS1 <- wrap_elements(full=FigureS1) / t1.effect.plot + plot_layout(heights = c(3, 1)) +  plot_annotation(tag_levels=list(c('', 'C')))

FigureS1 <- FigureS1 & theme(plot.tag = element_text(size=25))

ggsave(here::here("Figures", "FigureS1.png"), FigureS1, dpi=300, height=14, width=12)
ggsave(here::here("Figures", "FigureS1.svg"), FigureS1, dpi=300, height=14, width=12)

```



# Analysis of mid pregnancy sample
```{r}
T2Samples <- subset(FUT2Dataset, Timepoint == 2)

# 27 when everything is merged as is, double check
t2.aldex <- aldex.clr(t(T2Samples[, 36:dim(T2Samples)[2]]), T2Samples$`SecretorStatus`, mc.samples=1000, verbose=T)
t2.aldex.tt <- aldex.ttest(t2.aldex, paired.test=F, verbose=T)
t2.aldex.effect <- aldex.effect(t2.aldex, verbose=F)
t2.all <- data.frame(t2.aldex.tt, t2.aldex.effect, stringsAsFactors = F)
```

```{r}
par(mfrow=c(1,2))
aldex.plot(t2.all, type="MA", test="welch")

aldex.plot(t2.all, type="MW", test="welch")
t2.all$spp <- with(t2.all, reorder(rownames(t2.all), effect))
t2.all[["sign"]] = ifelse(t2.all[["effect"]] >= 0, "secretor", "non-secretor")

par(mfrow=c(1,1))

t2.all_topN <- t2.all[order(abs(t2.all$effect), decreasing=T), ][1:20, ]
t2.effect.plot <- ggplot(t2.all_topN, aes(x=spp, y=effect, fill=sign)) + geom_bar(stat="identity", width=0.5) + coord_flip() + theme_light() + scale_fill_discrete(name="Enriched in") + ylab('Effect') + xlab('Taxon') + theme(axis.text.y=element_text(size=12, face="italic"), axis.text.x = element_text(size=12), axis.title.x = element_text(size=14), axis.title.y = element_markdown(size=14))
t2.effect.plot
```

```{r}
FUT2Patients <- FUT2Dataset[!duplicated(FUT2Dataset$SubjectID), ]

MetadataFrame_Summary_SecretorStatus <- FUT2Patients %>% dplyr::select(SecretorStatus, LactobacillusDominant, CST, ethnicity.simplified)

# colnames(MetadataFrame_Summary) <- c('Secretor Status', 'Lactobacillus Dominant', 'CST')

summaryTable_S1 <- tbl_summary(MetadataFrame_Summary_SecretorStatus, by=c(`SecretorStatus`), statistic = all_continuous() ~ "{mean} ({sd})") 

```

