---
title: "FUT2 - Regression modelling analysis"
output: html_document
author:
  - Samit Kundu
  - Gonçalo Correia
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Parallel cross-sectional modelling

Modelling (GLMM) of the secretor data/vaginal microbiome data across three timepoints in pregnancy.

### 1.Modelling with Lactobacillus status at Timepoint 1 ("early" pregnancy)

Perform standard dummy coding regression using the GlmmTMB library to fit a random intercept model (ethnicity coded as a four level categorical variable [CSAsia, European, Afro-Caribbean and Other]) regressing gestational age (days) on the following covariates with a gamma distribution:

Maternal age (maternal.age)
BMI (bmi)
Previous Preterm birth or mid-trimester loss (fpbd):
  - no previous PTB or MTL
  - Previous PTB or MTL
Cervical cerclage (cerclage):
  - no cerclage
  - cerclage
Previous Lletz treatment(lletz):
  - no previous Lletz
  - previous Lletz treatment
Ethnicity (ethnicity):
  - White
  - Black
  - Asian
  - Arab
  - Mixed
  - Other
Secretor status (SecretorStatus):
  - Secretor
  - Non-secretor
Lactobacillus status (LactobacillusDominant in early, mid and late pregnancy):
  - L. Dominant - Lactobacillus dominated
  - L. Deplete - Lactobacillus depleted
CST (CST in early, mid and late pregnancy):
  - CST I
  - CST II
  - CST III
  - CST IV
  - CST V

ABO group:
  0 - A
  1 - AB
  2 - B
  3 - O

We apply a reflection transformation to the gestational age variable to use a gamma distribution ("lgest" variable).

Note
1. Lactobacillus status is defined using the VALENCIA classification algorithm output (L. Depleted = CST IV, L. Dominant = CST I, II, III, V)
2. Our previous analyses (not shown here) indicate only moderate evidence of multicollinearity (VIF values are all low, <10) between some variables (Previous PTB/MTL and previous Lletz).

# Package and data import
```{r}
library(glmmTMB)
library(car)
library(emmeans)
library(DHARMa)
library(readxl)
library(ggplot2)
library(readxl)
library(dplyr)
library(zCompositions)
require(magrittr)
library(ggtext)
library(gt)
library(gtsummary)
library(patchwork)


FUT2Dataset <- readxl::read_excel('./Data//FUT2-Dataset.xlsx', sheet='Clinical Data')

FUT2Dataset_Species <- readxl::read_excel('./Data//FUT2-Dataset.xlsx', sheet='Species Level Dataset')

# Remove 2 samples with very low counts to known species
FUT2Dataset <- subset(FUT2Dataset, !SampleID %in% c('T1A22'))
FUT2Dataset_Species <- subset(FUT2Dataset_Species, !SampleID %in% c('T1A22'))

colnames(FUT2Dataset)[colnames(FUT2Dataset) == 'Secretor Status'] <- "SecretorStatus"
FUT2Dataset$bmi <- as.numeric(FUT2Dataset$bmi)
FUT2Dataset$maternal.age <- as.numeric(FUT2Dataset$maternal.age)
FUT2Dataset$ABO <- factor(FUT2Dataset$ABO, levels=c('A', 'AB', 'B', 'O'))
FUT2Dataset$CST[grepl('IV', FUT2Dataset$CST)] <- 'IV'
FUT2Dataset$CST <- factor(FUT2Dataset$CST, levels=c('I', 'II', 'III', 'IV', 'V'))

predominantTaxa <- colnames(FUT2Dataset_Species[, 2:ncol(FUT2Dataset_Species)])[apply(FUT2Dataset_Species[, 2:ncol(FUT2Dataset_Species)], 1, FUN=function(x) { which.max(x)})]

FUT2Dataset$LactobacillusDominant_CST <- 'L. Dominant'
FUT2Dataset$LactobacillusDominant_CST[grepl('IV', FUT2Dataset$CST)] <- 'L. Depleted'
FUT2Dataset$LactobacillusDominant_CST[FUT2Dataset$subCST == 'IV-C3'] <- 'L. Dominant'
FUT2Dataset$LactobacillusDominant_CST[predominantTaxa %in% c('Lactobacillus', 'Lactobacillus delbrueckii', 'Bifidobacterium longum', 'Bifidobacterium breve', 
                                                             'Lactobacillus acidophilus', 'Lactobacillus rhamnosus', 'Limosilactobacillus')] <- 'L. Dominant'
FUT2Dataset$LactobacillusDominant_CST[predominantTaxa %in% c('Gardnerella uncultured bacterium', 'Gardnerella')] <- 'L. Depleted'

FUT2Dataset$LactobacillusDominant <- factor(FUT2Dataset$LactobacillusDominant, levels=c('L. Dominant', 'L. Depleted'))
FUT2Dataset$LactobacillusDominant_CST <- factor(FUT2Dataset$LactobacillusDominant_CST, levels=c('L. Dominant', 'L. Depleted'))
```

# Descriptive Statistics
## Figure 3 - Distribution of gestational length in days with secretor status and microbial composition
```{r}
CustomColours <- c('I'='forestgreen', 'II'='steelblue3', 'III'='darkorange3', 'IV'='firebrick3', 'V'='turquoise4')

# Timepoint 1 CST
plot_data_T1_CST <- subset(FUT2Dataset, Timepoint == 1) %>%
  mutate(group = interaction(SecretorStatus, CST, sep = "_"))

sample_sizes_T1_CST <- plot_data_T1_CST %>%
  group_by(SecretorStatus, CST) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(label = paste0(n))


SecretorGestAge_1_CST <- ggplot(plot_data_T1_CST, aes(x=SecretorStatus, fill=CST, y=delivery.gestation_total.days)) + geom_boxplot(alpha=0.8, outlier.shape = NA) + 
  geom_jitter(position=position_jitterdodge()) + scale_fill_manual(values=CustomColours) + 
  geom_text(data = sample_sizes_T1_CST, aes(x = SecretorStatus, y = 140, label = label, group = CST), 
            position = position_dodge(width = 0.75), vjust = 1.5,  size = 5) + theme_light() + 
  theme(legend.title = element_text(size=16),legend.text = element_text(size=15), axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 18), axis.title.x=element_text(size=18), axis.title.y=element_text(size=18)) + 
  xlab(NULL) + geom_hline(aes(yintercept=37*7), linetype='dashed', col='firebrick4') + ylab('Gestational length (days)') + ylim(c(140, 294)) 

# Timepoint 2 CST
plot_data_T2_CST <- subset(FUT2Dataset, Timepoint == 2) %>%
  mutate(group = interaction(SecretorStatus, CST, sep = "_"))

sample_sizes_T2_CST <- plot_data_T2_CST %>%
  group_by(SecretorStatus, CST) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(label = paste0(n))

SecretorGestAge_2_CST <- ggplot(plot_data_T2_CST, aes(x=SecretorStatus, fill=CST, y=delivery.gestation_total.days)) + geom_boxplot(alpha=0.8, outlier.shape = NA) + 
  geom_jitter(position=position_jitterdodge()) +  scale_fill_manual(values=CustomColours) +
  geom_text(data = sample_sizes_T2_CST, aes(x = SecretorStatus, y = 140, label = label, group = CST), 
            position = position_dodge(width = 0.75), vjust = 1.5,  size = 5) + theme_light() + 
   theme(legend.title = element_text(size=16),legend.text = element_text(size=15), axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 18), axis.title.x=element_text(size=18), axis.title.y=element_text(size=18))  + 
  xlab(NULL) + geom_hline(aes(yintercept=37*7), linetype='dashed', col='firebrick4') + ylab(NULL) + ylim(c(140, 294)) 

# Timepoint 3 CST
plot_data_T3_CST <- subset(FUT2Dataset, Timepoint == 3) %>%
  mutate(group = interaction(SecretorStatus, CST, sep = "_"))

sample_sizes_T3_CST <- plot_data_T3_CST %>%
  group_by(SecretorStatus, CST) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(label = paste0(n))

SecretorGestAge_3_CST <- ggplot(plot_data_T3_CST, aes(x=SecretorStatus, fill=CST, y=delivery.gestation_total.days)) + geom_boxplot(alpha=0.8, outlier.shape = NA) +
  geom_jitter(position=position_jitterdodge()) +  scale_fill_manual(values=CustomColours) + 
    geom_text(data = sample_sizes_T3_CST, aes(x = SecretorStatus, y = 140, label = label, group = CST), 
            position = position_dodge(width = 0.75), vjust = 1.5,  size = 5) + theme_light() + 
   theme(legend.title = element_text(size=16),legend.text = element_text(size=15), axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 18), axis.title.x=element_text(size=18), axis.title.y=element_text(size=18))   + 
  xlab(NULL)  + geom_hline(aes(yintercept=37*7), linetype='dashed', col='firebrick4') + ylab(NULL) + ylim(c(140, 294)) + ylab(NULL) 


# Lactobacillus dominant plots
CustomColours_LDom <- c('L. Dominant'='forestgreen', 'L. Depleted'='firebrick3')

# Timepoint 1 LDom vs LDepl
plot_data_T1 <- subset(FUT2Dataset, Timepoint == 1) %>%
  mutate(group = interaction(SecretorStatus, LactobacillusDominant, sep = "_"))

sample_sizes_T1 <- plot_data_T1 %>%
  group_by(SecretorStatus, LactobacillusDominant) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(label = paste0(n))

SecretorGestAge_1_LDom <- ggplot(plot_data_T1, aes(x=SecretorStatus, fill=LactobacillusDominant, y=delivery.gestation_total.days)) + geom_boxplot(alpha=0.8, outlier.shape = NA) + 
  geom_jitter(position=position_jitterdodge()) +  scale_fill_manual(values=CustomColours_LDom) + 
  geom_text(data = sample_sizes_T1, aes(x = SecretorStatus, y = 140, label = label, group = LactobacillusDominant), 
            position = position_dodge(width = 0.75), vjust = 1.5,  size = 5) + theme_light() + 
   theme(legend.title = element_text(size=16),legend.text = element_text(size=15), axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 18), axis.title.x=element_text(size=18), axis.title.y=element_text(size=18))  + 
  xlab(NULL)  + 
  geom_hline(aes(yintercept=37*7), linetype='dashed', col='firebrick4') + ylab('Gestational length (days)') + 
  guides(fill=guide_legend(title="*Lactobacillus* status"))   +
  theme(legend.title = element_markdown()) + ylim(c(140, 294))  + ggtitle('Early') + theme(plot.title = element_text(hjust = 0.5, size = 30, face = "plain"))  

# Timepoint 2 LDom vs LDepl
plot_data_T2 <- subset(FUT2Dataset, Timepoint == 2) %>%
  mutate(group = interaction(SecretorStatus, LactobacillusDominant, sep = "_"))

sample_sizes_T2 <- plot_data_T2 %>%
  group_by(SecretorStatus, LactobacillusDominant) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(label = paste0(n))

SecretorGestAge_2_LDom <- ggplot(plot_data_T2, aes(x=SecretorStatus, fill=LactobacillusDominant, y=delivery.gestation_total.days)) + geom_boxplot(alpha=0.8, outlier.shape = NA) + 
  geom_jitter(position=position_jitterdodge()) +  scale_fill_manual(values=CustomColours_LDom) + 
   geom_text(data = sample_sizes_T2, aes(x = SecretorStatus, y = 140, label = label, group = LactobacillusDominant), 
            position = position_dodge(width = 0.75), vjust = 1.5,  size = 5) + theme_light() + 
   theme(legend.title = element_text(size=16),legend.text = element_text(size=15), axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 18), axis.title.x=element_text(size=18), axis.title.y=element_text(size=18)) + 
  xlab(NULL) + geom_hline(aes(yintercept=37*7), linetype='dashed', col='firebrick4') + 
  ylab(NULL) + guides(fill=guide_legend(title="*Lactobacillus* status")) +  
  theme(legend.title = element_markdown()) + ylim(c(140, 294))  + ggtitle('Mid') +
    theme(plot.title = element_text(hjust = 0.5, size = 30, face = "plain"))

# Timepoint 3 LDom vs LDepl
plot_data_T3 <- subset(FUT2Dataset, Timepoint == 3) %>%
  mutate(group = interaction(SecretorStatus, LactobacillusDominant, sep = "_"))

sample_sizes_T3 <- plot_data_T3 %>%
  group_by(SecretorStatus, LactobacillusDominant) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(label = paste0(n))

SecretorGestAge_3_LDom <- ggplot(plot_data_T3, aes(x=SecretorStatus, fill=LactobacillusDominant, y=delivery.gestation_total.days)) + geom_boxplot(alpha=0.8, outlier.shape = NA) + 
  geom_jitter(position=position_jitterdodge()) +  scale_fill_manual(values=CustomColours_LDom) + 
  geom_text(data = sample_sizes_T3, aes(x = SecretorStatus, y = 140, label = label, group = LactobacillusDominant), 
            position = position_dodge(width = 0.75), vjust = 1.5,  size = 5) +  theme_light() + 
   theme(legend.title = element_text(size=16),legend.text = element_text(size=15), axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 18), axis.title.x=element_text(size=18), axis.title.y=element_text(size=18))  + 
  xlab(NULL) + geom_hline(aes(yintercept=37*7), linetype='dashed', col='firebrick4') + 
  ylab(NULL)  + guides(fill=guide_legend(title="*Lactobacillus* status")) + 
  theme(legend.title = element_markdown()) + ylim(c(140, 294))  + ggtitle('Late') +
    theme(plot.title = element_text(hjust = 0.5, size = 30, face = "plain"))

Figure3_LD <- (SecretorGestAge_1_LDom | SecretorGestAge_2_LDom | SecretorGestAge_3_LDom)
Figure3_CST <- (SecretorGestAge_1_CST | SecretorGestAge_2_CST | SecretorGestAge_3_CST)

Figure3_CST <- Figure3_CST + plot_layout(guides='collect')
Figure3_LD <- Figure3_LD + plot_layout(guides='collect') 

Figure3 <- (Figure3_LD / Figure3_CST)  & theme(legend.justification = "left")
Figure3 <- Figure3  + plot_annotation(tag_levels=list(c('A', '', '', 'B', '', ''))) 
Figure3 <- Figure3 & theme(plot.tag = element_text(size=25))

ggsave(here::here('Figures', 'Figure3.png'), dpi=300, width=18, height=20)
ggsave(here::here('Figures', 'Figure3.svg'), dpi=300, width=18, height=20)
```

## Table S1
```{r}
MetadataFrame_PerPregnancy <- subset(FUT2Dataset, !duplicated(SubjectID))
MetadataFrame_PerPregnancy$BMICat <- cut(MetadataFrame_PerPregnancy$bmi, 
                   breaks=c(-Inf, 18.5, 24.99, 29.99, Inf), 
                   labels=c("<18.5","18.5 - 24.99","25-29.99", ">30"))

MetadataFrame_PerPregnancy$OutcomeCat <- cut(MetadataFrame_PerPregnancy$delivery.gestation_total.days, 
                   breaks=c(-Inf, 196, 224, 259, Inf), 
                   labels=c("Extreme Preterm","Very Preterm", "Moderate to late Preterm", "Term"), right=FALSE)


MetadataFrame_PerPregnancy$cerclage <- factor(MetadataFrame_PerPregnancy$cerclage)
MetadataFrame_PerPregnancy$lletz <- factor(MetadataFrame_PerPregnancy$lletz)

MetadataFrame_PerPregnancy$Outcome <- MetadataFrame_PerPregnancy$delivery.gestation_total.days < 259

MetadataFrame_Summary <- MetadataFrame_PerPregnancy %>% dplyr::select(delivery.gestation_total.days, OutcomeCat, maternal.age, SecretorStatus, BMICat, ethnicity.simplified, lletz, cerclage, previous.ptb)

MetadataFrame_Summary$ethnicity.simplified[MetadataFrame_Summary$ethnicity.simplified == 'Mixed black/white'] <- 'Mixed/Other'
MetadataFrame_Summary$ethnicity.simplified[MetadataFrame_Summary$ethnicity.simplified == 'Mixed'] <- 'Mixed/Other'
MetadataFrame_Summary$ethnicity.simplified[MetadataFrame_Summary$ethnicity.simplified == 'Other'] <- 'Mixed/Other'

# colnames(MetadataFrame_Summary) <- c('Secretor Status', 'Lactobacillus Dominant', 'CST')

summaryTable_S1 <- tbl_summary(MetadataFrame_Summary, by=c(`SecretorStatus`), statistic = all_continuous() ~ "{mean} ({sd})") 
# %>% modify_header(label = '**Outcome**') %>% add_p(pvalue_fun = function(x) style_pvalue(x, digits = 3))
```

```{r}
Timepoint_Distribution <- ggplot(FUT2Dataset, aes(x=sampling.gestation_total.days, fill=factor(Timepoint))) + geom_histogram() + geom_vline(xintercept = 133.0, linetype='dashed') +
  geom_vline(xintercept = 178.0, linetype='dashed')  + xlab('Gestational age at sampling (days)') + theme_light() + scale_fill_discrete(name='Time point', labels=c('1 - Early', '2 - Mid', '3 - Late'))

SamplesPerPatient <- FUT2Dataset %>% group_by(SubjectID) %>% mutate(n=n()) %>% select(n, SubjectID, SecretorStatus)  %>% distinct(SubjectID, .keep_all = TRUE) 

table(SamplesPerPatient$n, SamplesPerPatient$SecretorStatus)
```
# CST distribution per timepoint
```{r}
TP1 <- subset(FUT2Dataset, Timepoint == 1)
TP2 <- subset(FUT2Dataset, Timepoint == 2)
TP3 <- subset(FUT2Dataset, Timepoint == 3)

table(TP1$CST)/dim(TP1)[1]
table(TP2$CST)/dim(TP2)[1]
table(TP3$CST)/dim(TP3)[1]
```

# Distribution of outcome at each timepoint
```{r}
MetadataFrame_OutcomeTimepoint <-FUT2Dataset

MetadataFrame_OutcomeTimepoint$OutcomeCat <- cut(MetadataFrame_OutcomeTimepoint$delivery.gestation_total.days, 
                   breaks=c(-Inf, 196, 224, 259, Inf), 
                   labels=c("Extreme Preterm","Very Preterm", "Moderate to late Preterm", "Term"), right=FALSE)

MetadataFrame_Summary <- MetadataFrame_OutcomeTimepoint %>% dplyr::select(OutcomeCat, SecretorStatus, Timepoint)

# colnames(MetadataFrame_Summary) <- c('Secretor Status', 'Lactobacillus Dominant', 'CST')

summaryTable_TimepointOutcome <-  MetadataFrame_Summary %>% tbl_strata(
    strata = SecretorStatus,
    ~.x %>%  tbl_summary(
        by = Timepoint))
```


## Genotype statistics
```{r}
genotypes <- MetadataFrame_PerPregnancy[, c('SubjectID', 'SecretorStatus', 'ethnicity.simplified', 'I129F', 'P101L', 'R191X', 'W143X')]

# heterozygous (secretor or not)
heterozygous <- genotypes %>% filter(I129F == 'F|I' | P101L == 'L|P' | R191X %in% c("X|R", "R|Q") | W143X == 'W|X')
# non-secretor homozygous genotypes
nonSecretorhomozygous <-  genotypes %>% filter(I129F == 'F|F' | P101L == 'L|L' | W143X == 'X|X')

# Proportion of 'X|X' se428 in non-secretor homozygous 
sum(nonSecretorhomozygous$W143X =='X|X')/nrow(nonSecretorhomozygous) * 100

# Proportion of se428 heterozygous
sum(heterozygous$W143X =='W|X')/nrow(heterozygous) * 100
# Ethnic breakdown of genotypes
table(genotypes$I129F, genotypes$ethnicity.simplified)
table(genotypes$P101L, genotypes$ethnicity.simplified)
table(genotypes$W143X, genotypes$ethnicity.simplified)
```

## Info used for table S2
```{r}
countsT1 <- subset(FUT2Dataset, Timepoint == 1) %>% select(delivery.gestation_total.days, SecretorStatus, LactobacillusDominant, CST) %>% group_by(SecretorStatus, LactobacillusDominant) %>% summarise(values = paste0(format(mean(delivery.gestation_total.days), nsmall=2), ' (', n(), ')')) %>% gt()

countsT2 <- subset(FUT2Dataset, Timepoint == 2) %>% select(delivery.gestation_total.days, SecretorStatus, LactobacillusDominant, CST) %>% group_by(SecretorStatus, LactobacillusDominant) %>% summarise(values = paste0(format(mean(delivery.gestation_total.days), nsmall=2), ' (', n(), ')')) %>% gt()

countsT3 <- subset(FUT2Dataset, Timepoint == 3) %>% select(delivery.gestation_total.days, SecretorStatus, LactobacillusDominant, CST) %>% group_by(SecretorStatus, LactobacillusDominant) %>% summarise(values = paste0(format(mean(delivery.gestation_total.days), nsmall=2), ' (', n(), ')')) %>% gt()

countsT1_CST <- subset(FUT2Dataset, Timepoint == 1) %>% select(delivery.gestation_total.days, SecretorStatus, LactobacillusDominant, CST) %>% group_by(SecretorStatus, CST) %>% summarise(values = paste0(format(mean(delivery.gestation_total.days), nsmall=2), ' (', n(), ')')) %>% gt()

countsT2_CST <- subset(FUT2Dataset, Timepoint == 2) %>% select(delivery.gestation_total.days, SecretorStatus, LactobacillusDominant, CST) %>% group_by(SecretorStatus, CST) %>% summarise(values = paste0(format(mean(delivery.gestation_total.days), nsmall=2), ' (', n(), ')')) %>% gt()

countsT3_CST <- subset(FUT2Dataset, Timepoint == 3) %>% select(delivery.gestation_total.days, SecretorStatus, LactobacillusDominant, CST) %>% group_by(SecretorStatus, CST) %>% summarise(values = paste0(format(mean(delivery.gestation_total.days), nsmall=2), ' (', n(), ')')) %>% gt()
```
# Descriptive statistics and comparisons of delivery gestation
```{r}
Table_Mean_SD_Ldom <- subset(FUT2Dataset, Timepoint == 1) %>% select(delivery.gestation_total.days, SecretorStatus, LactobacillusDominant, CST) %>% group_by(SecretorStatus, LactobacillusDominant) %>% summarise(values = paste0(format(mean(delivery.gestation_total.days), nsmall=2), ' (', sd(delivery.gestation_total.days), ')')) %>% gt()

# Get the tukey adjusted p-values for the sub-comparisons
pvalues_TP1_LDominant <- pairs(emmeans(lm(formula = delivery.gestation_total.days ~ 1 + LactobacillusDominant*SecretorStatus, data=FUT2_TP1), specs=~LactobacillusDominant*SecretorStatus))

Table_Mean_SD_CST <- subset(FUT2Dataset, Timepoint == 1) %>% select(delivery.gestation_total.days, SecretorStatus, LactobacillusDominant, CST) %>% group_by(SecretorStatus, CST) %>% summarise(values = paste0(format(mean(delivery.gestation_total.days), nsmall=2), ' (', sd(delivery.gestation_total.days), ')')) %>% gt()

pvalues_TP1_CST <- pairs(emmeans(lm(formula = delivery.gestation_total.days ~ 1 + CST*SecretorStatus, data=FUT2_TP1), specs=~CST*SecretorStatus))

pvalues_TP1_CST_Noadj <- pairs(emmeans(lm(formula = delivery.gestation_total.days ~ 1 + CST*SecretorStatus, data=FUT2_TP1), specs=~CST*SecretorStatus), adjust='none')


```


# Regression Analysis
## Early (TP1) Analysis
```{r}
FUT2_TP1 <- subset(FUT2Dataset, Timepoint == 1)

# reflectorTransformation 
# FUT2_TP1$LGest <- FUT2_TP1$delivery.gestation_total.days
FUT2_TP1$LGest <- (max(FUT2_TP1$delivery.gestation_total.days) + 1 ) - FUT2_TP1$`delivery.gestation_total.days`
FUT2_TP1 <- subset(FUT2_TP1, previous.ptb != 'NA')
FUT2_TP1$previous.ptb <- factor(FUT2_TP1$previous.ptb)
# Re-level factor to have non-secretor on model coefs by default
FUT2_TP1$SecretorStatus <- factor(FUT2_TP1$SecretorStatus, levels=c('Secretor', 'Non-secretor'))

#Run GLMM incoroporating Lactobacillus status at Timepoint 1 and secretor status - Original Model
# glmm_t1 <- glmmTMB(lgest~Age+fbmi+fpbd+fstitch+floop+fsec+ft1ld+fsec:ft1ld+(1|feth), family=Gamma(link="log"), data=FUT2_TP1)

glmm_t1 <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2_TP1)

#Model diagnostics
simglmmT1 <- simulateResiduals(fittedModel=glmm_t1, plot=T)
testDispersion(simglmmT1)

#Perform analysis of deviance using the Anova (using Anova from the car package rather than the core anova function as the former automatically takes care of unbalanced designs)
Anova(glmm_t1)
summary(glmm_t1)

# Model using CST instead of LDepleted status
glmm_t1_CST <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz + SecretorStatus + CST + SecretorStatus:CST + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2_TP1)

#Model diagnostics
simglmmT1 <- simulateResiduals(fittedModel=glmm_t1_CST, plot=T)
testDispersion(simglmmT1)

#Perform analysis of deviance using the Anova (using Anova from the car package rather than the core anova function as the former automatically takes care of unbalanced designs)
Anova(glmm_t1_CST)
summary(glmm_t1_CST)

results_table_t1 <-
  Anova(glmm_t1) %>% magrittr::set_rownames(c('Maternal age', 'BMI', "Previous PTB", 'Cervical cerclage', 'Previous cervical excisional treatment', 'Secretor Status', 'Lactobacillus Dominant', 'Secretor Status : Lactobacillus Dominant')) %>% 
  gt(rownames_to_stub = T) %>%
  fmt_number( columns = c(2),
    decimals = 3,
    suffixing = F)  %>% 
  fmt_scientific( columns = c(4),
    decimals = 3) 

results_table_t1_CST <-
  Anova(glmm_t1_CST) %>% magrittr::set_rownames(c('Maternal age', 'BMI', "Previous PTB", 'Cervical cerclage', 'Previous cervical excisional treatment', 'Secretor Status', 'CST', 'Secretor Status : CST')) %>% 
  gt(rownames_to_stub = T) %>%
  fmt_number( columns = c(2),
    decimals = 3,
    suffixing = F)  %>% 
  fmt_scientific( columns = c(4),
    decimals = 3) 

```
# Model estimates - Early
```{r}
# Part of Table 1
ciEstimates_T1 <- confint(glmm_t1)

estimatesTable_table_t1 <- 
  tibble::as_tibble(ciEstimates_T1) %>% mutate(coefficient=rownames(ciEstimates_T1)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 

ciEstimates_T1_CST <- confint(glmm_t1_CST)

estimatesTable_table_t1_cst <- 
  tibble::as_tibble(ciEstimates_T1_CST) %>% mutate(coefficient=rownames(ciEstimates_T1_CST)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 


emm1 <- emmeans(glmm_t1, specs = ~SecretorStatus:LactobacillusDominant, type="response")
emm1_CST <- emmeans(glmm_t1_CST, specs = ~SecretorStatus:CST, type="response")

emm1_ctr <- contrast(emm1, method="pairwise", adjust = "none")
emm1_ctr_CST <- contrast(emm1_CST, method="pairwise", adjust = "none")
```

The analysis of deviance results indicate that Previous PTB/MTL, Cerclage, Previous Lletz and the secretor status:Lactobacillus interaction terms are significant covariates. Residual analysis using the Dharma package indicates that the model fits well for each covariate. The coefficients indicate that the non-secretors with a Lactobacillus depleted microbiome have shorter gestational lengths (hence positive relationship with the inverted variable) compared with the baseline (women who are secretors with Lactobacillus dominated microbiomes). Women who have had a previous PTB/MTL also have shorter gestational lengths (compared with women who have not) as do women who get a cervical stitch (cerclage). 

Post hoc comparison using the emmeans package suggests that the most significant comparison is between the non-secretors with Lactobacillus depleted microbiomes and non-secretors with Lactobacillus dominant microbiomes (comparison of secretors that have Lactobacillus depleted microbiomes with non-secretors that also have Lactobacillus deplete microbiomes is close to significance).

Adjustment for Blood Group
```{r}
# Fit a similar model with adjustment for ABO blood group
# Results shown in Table S6
glmm_t1_abo <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  ABO + SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2_TP1)


ciEstimates_T1_abo <- confint(glmm_t1_abo)

estimatesTable_table_t1_abo <- 
  tibble::as_tibble(ciEstimates_T1_abo) %>% mutate(coefficient=rownames(ciEstimates_T1_abo)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 

```


## Mid (TP2) Analysis
```{r}
FUT2_TP2 <- subset(FUT2Dataset, Timepoint == 2)

# reflectorTransformation 
FUT2_TP2$LGest <- (max(FUT2_TP2$delivery.gestation_total.days) + 1 ) - FUT2_TP2$`delivery.gestation_total.days`
FUT2_TP2 <- subset(FUT2_TP2, previous.ptb != 'NA')
FUT2_TP2$previous.ptb <- factor(FUT2_TP2$previous.ptb)
# Re-level factor to have non-secretor on model coefs by default
FUT2_TP2$SecretorStatus <- factor(FUT2_TP2$SecretorStatus, levels=c('Secretor', 'Non-secretor'))

glmm_t2 <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2_TP2)

#Model diagnostics
simglmmT2 <- simulateResiduals(fittedModel=glmm_t2, plot=T)
testDispersion(simglmmT2)

#Perform analysis of deviance using the Anova (using Anova from the car package rather than the core anova function as the former automatically takes care of unbalanced designs)
Anova(glmm_t2)
summary(glmm_t2)

# Model using CST instead of LDepleted status
glmm_t2_CST <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz + SecretorStatus + CST + SecretorStatus:CST + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2_TP2)

#Model diagnostics
simglmmT2 <- simulateResiduals(fittedModel=glmm_t2_CST, plot=T)
testDispersion(simglmmT2)

#Perform analysis of deviance using the Anova (using Anova from the car package rather than the core anova function as the former automatically takes care of unbalanced designs)
Anova(glmm_t2_CST)
summary(glmm_t2_CST)

results_table_t2 <-
  Anova(glmm_t2) %>% magrittr::set_rownames(c('Maternal age', 'BMI', "Previous PTB", 'Cervical cerclage', 'Previous cervical excisional treatment', 'Secretor Status', 'Lactobacillus Dominant', 'Secretor Status : Lactobacillus Dominant')) %>% 
  gt(rownames_to_stub = T) %>%
  fmt_number( columns = c(2),
    decimals = 3,
    suffixing = F)  %>% 
  fmt_scientific( columns = c(4),
    decimals = 3) 

results_table_t2_CST <-
  Anova(glmm_t2_CST) %>% magrittr::set_rownames(c('Maternal age', 'BMI', "Previous PTB", 'Cervical cerclage', 'Previous cervical excisional treatment', 'Secretor Status', 'CST', 'Secretor Status : CST')) %>% 
  gt(rownames_to_stub = T) %>%
  fmt_number( columns = c(2),
    decimals = 3,
    suffixing = F)  %>% 
  fmt_scientific( columns = c(4),
    decimals = 3) 

```

## Model coefficients - Mid
```{r}
# Part of Table 1
ciEstimates_T2 <- confint(glmm_t2)

estimatesTable_table_t2 <- 
  tibble::as_tibble(ciEstimates_T2) %>% mutate(coefficient=rownames(ciEstimates_T2)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 

ciEstimates_T2_CST <- confint(glmm_t2_CST)

estimatesTable_table_t2_cst <- 
  tibble::as_tibble(ciEstimates_T2_CST) %>% mutate(coefficient=rownames(ciEstimates_T2_CST)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 

```

Adjustment for Blood Group
```{r}
# Fit a similar model with adjustment for ABO blood group
# Results shown in Table S5
glmm_t2_abo <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  ABO + SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2_TP2)

ciEstimates_T2_abo <- confint(glmm_t2_abo)

estimatesTable_table_t2_abo <- 
  tibble::as_tibble(ciEstimates_T2_abo) %>% mutate(coefficient=rownames(ciEstimates_T2_abo)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 
```


## Late (TP3) Analysis
```{r}
FUT2_TP3 <- subset(FUT2Dataset, Timepoint == 3)

# reflectorTransformation 
FUT2_TP3$LGest <- (max(FUT2_TP3$delivery.gestation_total.days) + 1 ) - FUT2_TP3$`delivery.gestation_total.days`
FUT2_TP3 <- subset(FUT2_TP3, previous.ptb != 'NA')
FUT2_TP3$previous.ptb <- factor(FUT2_TP3$previous.ptb)
# Re-level factor to have non-secretor on model coefs by default
FUT2_TP3$SecretorStatus <- factor(FUT2_TP3$SecretorStatus, levels=c('Secretor', 'Non-secretor'))

#Run GLMM incoroporating Lactobacillus status at Timepoint 1 and secretor status - Original Model
glmm_t3 <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2_TP3)

#Model diagnostics
simglmmT3 <- simulateResiduals(fittedModel=glmm_t3, plot=T)
testDispersion(simglmmT3)

#Perform analysis of deviance using the Anova (using Anova from the car package rather than the core anova function as the former automatically takes care of unbalanced designs)
Anova(glmm_t3)
summary(glmm_t3)

# Model using CST instead of LDepleted status
glmm_t3_CST <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz + SecretorStatus + CST + SecretorStatus:CST + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2_TP3)

#Model diagnostics
simglmmT3 <- simulateResiduals(fittedModel=glmm_t3_CST, plot=T)
testDispersion(simglmmT3)

#Perform analysis of deviance using the Anova (using Anova from the car package rather than the core anova function as the former automatically takes care of unbalanced designs)
Anova(glmm_t3_CST)
summary(glmm_t3_CST)

results_table_t3 <-
  Anova(glmm_t3) %>% magrittr::set_rownames(c('Maternal age', 'BMI', "Previous PTB", 'Cervical cerclage', 'Previous cervical excisional treatment', 'Secretor Status', 'Lactobacillus Dominant', 'Secretor Status : Lactobacillus Dominant')) %>% 
  gt(rownames_to_stub = T) %>%
  fmt_number( columns = c(2),
    decimals = 3,
    suffixing = F)  %>% 
  fmt_scientific( columns = c(4),
    decimals = 3) 

results_table_t3_CST <-
  Anova(glmm_t3_CST) %>% magrittr::set_rownames(c('Maternal age', 'BMI', "Previous PTB", 'Cervical cerclage', 'Previous cervical excisional treatment', 'Secretor Status', 'CST', 'Secretor Status : CST')) %>% 
  gt(rownames_to_stub = T) %>%
  fmt_number( columns = c(2),
    decimals = 3,
    suffixing = F)  %>% 
  fmt_scientific( columns = c(4),
    decimals = 3) 
```
## Model coefficients - Late
```{r}
# Part of Table 1
ciEstimates_T3 <- confint(glmm_t3)

estimatesTable_table_t3 <- 
  tibble::as_tibble(ciEstimates_T3) %>% mutate(coefficient=rownames(ciEstimates_T3)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 

ciEstimates_T3_CST <- confint(glmm_t3_CST)

estimatesTable_table_t3_cst <- 
  tibble::as_tibble(ciEstimates_T3_CST) %>% mutate(coefficient=rownames(ciEstimates_T3_CST)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 

```

Adjustment for Blood Group
```{r}
# Fit a similar model with adjustment for ABO blood group
# Results shown in Table S5
glmm_t3_abo <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  ABO + SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2_TP3)

ciEstimates_T3_abo <- confint(glmm_t3_abo)

estimatesTable_table_t3_abo <- 
  tibble::as_tibble(ciEstimates_T3_abo) %>% mutate(coefficient=rownames(ciEstimates_T3_abo)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 
```

#Perform a simpler GLM (for comparison with the GLMM)
```{r}
glm_t1 <- glmmTMB(LGest ~ maternal.age + bmi  + previous.ptb + cerclage + lletz +  SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant, family=Gamma(link="log"), data=FUT2_TP1)

AIC(glm_t1,glmm_t1)
```
Comparing the random intercept model with the GLM indicates that the latter is actually a slightly better fit (lower AIC), i.e., ethnicity does not appear to be a major factor (bias towards European women in this cohort may explain lack of signal with respect to ethnicity).
```{r}
simglmmT1 <- simulateResiduals(fittedModel=glmm_t1, plot=T)
testDispersion(simglmmT1)
emm1 <- emmeans(glmm_t1, specs = ~SecretorStatus:LactobacillusDominant, type="response")
emm1
contrast(emm1, method="pairwise", adjust = "none")
```

# Sensitivity analysis of Timepoint 1 results for Lactobacillus dominant vs Depleted and secretor status
```{r, warning=F}
FUT2Dataset_Genera <- readxl::read_excel('./Data//FUT2-Dataset.xlsx', sheet='Genera Level Dataset')
FUT2Dataset_Genera <- merge(FUT2Dataset, FUT2Dataset_Genera, by='SampleID')

FUT2Dataset_Genera_TP1 <- subset(FUT2Dataset_Genera, Timepoint == 1)

# reflectorTransformation 
# FUT2_TP1$LGest <- FUT2_TP1$delivery.gestation_total.days
FUT2Dataset_Genera_TP1$LGest <- (max(FUT2Dataset_Genera_TP1$delivery.gestation_total.days) + 1) - FUT2Dataset_Genera_TP1$`delivery.gestation_total.days`
FUT2Dataset_Genera_TP1 <- subset(FUT2Dataset_Genera_TP1, previous.ptb != 'NA')
FUT2Dataset_Genera_TP1$previous.ptb <- factor(FUT2Dataset_Genera_TP1$previous.ptb)
# Re-level factor to have non-secretor on model coefs by default
FUT2Dataset_Genera_TP1$SecretorStatus <- factor(FUT2Dataset_Genera_TP1$SecretorStatus, levels=c('Secretor', 'Non-secretor'))

glmm_t1_9 <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2Dataset_Genera_TP1)

#Run GLMM incoroporating Lactobacillus status at Timepoint 1 and secretor status - Original Model
FUT2Dataset_Genera_TP1$LactobacillusDominant[FUT2Dataset_Genera_TP1$Lactobacillus < 0.75] <- 'L. Depleted'
FUT2Dataset_Genera_TP1$LactobacillusDominant[FUT2Dataset_Genera_TP1$Lactobacillus > 0.75] <- 'L. Dominant'

glmm_t1_75 <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2Dataset_Genera_TP1)

Anova(glmm_t1_75)
summary(glmm_t1_75)

FUT2Dataset_Genera_TP1$LactobacillusDominant[FUT2Dataset_Genera_TP1$Lactobacillus < 0.5] <- 'L. Depleted'
FUT2Dataset_Genera_TP1$LactobacillusDominant[FUT2Dataset_Genera_TP1$Lactobacillus > 0.5] <- 'L. Dominant'

glmm_t1_50 <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2Dataset_Genera_TP1)

Anova(glmm_t1_50)
summary(glmm_t1_50)


FUT2Dataset_Genera_TP1$LactobacillusDominant[FUT2Dataset_Genera_TP1$Lactobacillus < 0.9] <- 'L. Depleted'
FUT2Dataset_Genera_TP1$LactobacillusDominant[FUT2Dataset_Genera_TP1$Lactobacillus > 0.9] <- 'L. Dominant'

pvals <- c()

thresholds <- seq(0.9, 0.25, -0.02)

for (threshold in thresholds) {

  
FUT2Dataset_Genera_TP1$LactobacillusDominant[FUT2Dataset_Genera_TP1$Lactobacillus < threshold] <- 'L. Depleted'
FUT2Dataset_Genera_TP1$LactobacillusDominant[FUT2Dataset_Genera_TP1$Lactobacillus > threshold] <- 'L. Dominant'

glmm_t1_sens <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  SecretorStatus + LactobacillusDominant + SecretorStatus:LactobacillusDominant + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2Dataset_Genera_TP1)

AnovaTable <- Anova(glmm_t1_sens)
pvals <- c(pvals, AnovaTable$`Pr(>Chisq)`[8])
}

sensResults <- data.frame(cbind(LDomvsLDeplThreshold=thresholds, pvalue=pvals))

sensitivityAnalysisPlot <- ggplot(sensResults, aes(x=LDomvsLDeplThreshold, y=pvalue)) + geom_line(color='steelblue4', size=0.5) + geom_point(color='steelblue4', size=2) + theme_light() + geom_hline(yintercept=0.05, linetype='dashed', color='firebrick4') + ylab('*Lactobacillus* Dominant x Secretor Status Interaction p-value') + 
  xlab("*Lactobacillus* relative abundance threshold") +
  theme(axis.title.x = element_markdown(),axis.title.y = element_markdown())

# ggsave(here::here('Figures', 'LactoDomvsDeplSensitivityAnalysis.png'), sensitivityAnalysisPlot, dpi=300, width=8, height=8)

```
```{r}
FUT2Dataset_Genera <- readxl::read_excel('./Data//FUT2-Dataset.xlsx', sheet='Genera Level Dataset')
FUT2Dataset_Genera <- merge(FUT2Dataset, FUT2Dataset_Genera, by='SampleID')

FUT2Dataset_Genera_TP1 <- subset(FUT2Dataset_Genera, Timepoint == 1)

# reflectorTransformation 
# FUT2_TP1$LGest <- FUT2_TP1$delivery.gestation_total.days
FUT2Dataset_Genera_TP1$LGest <- (max(FUT2Dataset_Genera_TP1$delivery.gestation_total.days) + 1) - FUT2Dataset_Genera_TP1$`delivery.gestation_total.days`
FUT2Dataset_Genera_TP1 <- subset(FUT2Dataset_Genera_TP1, previous.ptb != 'NA')
FUT2Dataset_Genera_TP1$previous.ptb <- factor(FUT2Dataset_Genera_TP1$previous.ptb)
# Re-level factor to have non-secretor on model coefs by default
FUT2Dataset_Genera_TP1$SecretorStatus <- factor(FUT2Dataset_Genera_TP1$SecretorStatus, levels=c('Secretor', 'Non-secretor'))


glmm_t1_LactoRel <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  SecretorStatus + Lactobacillus + SecretorStatus:Lactobacillus + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2Dataset_Genera_TP1)

Anova(glmm_t1_LactoRel)

results_table_t1_LactoRel <-
  Anova(glmm_t1_LactoRel) %>% magrittr::set_rownames(c('Maternal age', 'BMI', "Previous PTB", 'Cervical cerclage', 'Previous cervical excisional treatment', 'Secretor Status', 'Lactobacillus', 'Secretor Status : Lactobacillus')) %>% 
  gt(rownames_to_stub = T) %>%
  fmt_number( columns = c(2),
    decimals = 3,
    suffixing = F)  %>% 
  fmt_scientific( columns = c(4),
    decimals = 3) 

glmm_t1_LDomCST <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  SecretorStatus + LactobacillusDominant_CST + SecretorStatus:LactobacillusDominant_CST + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2Dataset_Genera_TP1)

Anova(glmm_t1_LDomCST)

results_table_t1_LDomCST <-
  Anova(glmm_t1_LDomCST) %>% magrittr::set_rownames(c('Maternal age', 'BMI', "Previous PTB", 'Cervical cerclage', 'Previous cervical excisional treatment', 'Secretor Status', 'Lactobacillus Dominant', 'Secretor Status : Lactobacillus Dominant')) %>% 
  gt(rownames_to_stub = T) %>%
  fmt_number( columns = c(2),
    decimals = 3,
    suffixing = F)  %>% 
  fmt_scientific( columns = c(4),
    decimals = 3) 

# Use Ldepl = IV, without accounting for VALENCIA pooling other Lacto and B. breve as part of IV
FUT2Dataset_Genera_TP1$LactobacillusDominant_CST <- 'L. Dominant'
FUT2Dataset_Genera_TP1$LactobacillusDominant_CST[grepl('IV', FUT2Dataset_Genera_TP1$CST)] <- 'L. Depleted'
FUT2Dataset_Genera_TP1$LactobacillusDominant_CST <- factor(FUT2Dataset_Genera_TP1$LactobacillusDominant_CST, levels=c('L. Dominant', 'L. Depleted'))

glmm_t1_LDomCST_IvsIVClassic <- glmmTMB(LGest ~ maternal.age + bmi + previous.ptb + cerclage + lletz +  SecretorStatus + LactobacillusDominant_CST + SecretorStatus:LactobacillusDominant_CST + (1|ethnicity.simplified), family=Gamma(link="log"), data=FUT2Dataset_Genera_TP1)

Anova(glmm_t1_LDomCST_IvsIVClassic)

results_table_t1_LDomCST_IvsIVClassic <-
  Anova(glmm_t1_LDomCST_IvsIVClassic) %>% magrittr::set_rownames(c('Maternal age', 'BMI', "Previous PTB", 'Cervical cerclage', 'Previous cervical excisional treatment', 'Secretor Status', 'Lactobacillus Dominant', 'Secretor Status : Lactobacillus Dominant')) %>% 
  gt(rownames_to_stub = T) %>%
  fmt_number( columns = c(2),
    decimals = 3,
    suffixing = F)  %>% 
  fmt_scientific( columns = c(4),
    decimals = 3) 

```

```{r}
ciEstimates_T1_RelAbundance <- confint(glmm_t1_LactoRel)

estimatesTable_table_t1_RelAbundance <- 
  tibble::as_tibble(ciEstimates_T1_RelAbundance) %>% mutate(coefficient=rownames(ciEstimates_T1_RelAbundance)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 

ciEstimates_T1_LDomCST <- confint(glmm_t1_LDomCST)

estimatesTable_table_t1_LDomCST <- 
  tibble::as_tibble(ciEstimates_T1_LDomCST) %>% mutate(coefficient=rownames(ciEstimates_T1_LDomCST)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 


ciEstimates_T1_LDomCST_IvsIVClassic <- confint(glmm_t1_LDomCST_IvsIVClassic)

estimatesTable_table_t1_LDomCST_IvsIVClassic <- 
  tibble::as_tibble(ciEstimates_T1_LDomCST_IvsIVClassic) %>% mutate(coefficient=rownames(ciEstimates_T1_LDomCST_IvsIVClassic)) %>% gt(rowname_col = 'coefficient') %>%
  fmt_number(columns = c(1, 2, 3),
    decimals = 3,
    suffixing = F) 


```